---
title: "Banana Sweetness Analysis"
authors:
  - name: Maeve McGillycuddy
format: 
  html:
    toc: true
    code-fold: true
    smooth-scroll: true
    embed-resources: true
---

## Background

We are making banana bread at work today and want the best possible ingredients. 
A key question is whether bananas from different **regions** have different 
**sweetness**, since sugar content (measured in degrees Brix) affects the 
flavour and texture of the final loaf.

This report uses the [Banana Quality data](https://www.kaggle.com/datasets/mrmars1010/banana-quality-dataset) from Kaggle to investigate whether `region` is associated with `sugar_content_brix`.


## Cleaning data

The raw data is loaded and cleaned using `01_clean.R`. This script standardises 
column names, parses dates, coerces categorical variables to factors, and checks 
for missing or implausible values.

```{r, results = 'hide', message=F, warning=F}
source(here::here("src", "01_clean.R"))
```


## Exploration 

We first visualise the relationship between region and sugar 
content to see  the differences between groups.

```{r}
bananas %>% 
ggplot(aes(x = reorder(region, sugar_content_brix, median),
                    y = sugar_content_brix, fill = region)) +
  geom_violin(alpha = 0.5, trim = FALSE) +
  geom_point(position = position_jitter(width = 0.15, seed = 42),
           alpha = 0.2, size = 1.0) +
  geom_boxplot( alpha = 0.2) +
  labs(x = "Region", y = "Sugar Content (Brix)",
       title = "Sugar content of bananas from differnt regions") +
  theme_classic(base_size = 13) + 
  theme(legend.position = "none") +
  scale_fill_brewer(palette = "Set3")
```

## Linear regression

To examine the relationship between sugar content and region, we fit a 
simple linear regression to `sugar_content_brix` against `region` 
as the predictor. 

```{r, message=F, warning=F}
model <- lm(sugar_content_brix ~ region, data = bananas)
```

The estimated mean sugar content for each region is shown below, along with their 
95% confidence intervals.

```{r, message=F, warning=F}
library(emmeans)
library(broom)

emmeans(model, ~ region) |>
  tidy(conf.int = TRUE) |>
  ggplot(aes(x = estimate, y = reorder(region, estimate))) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high)) +
  labs(x = "Estimated mean Sugar Content (Brix)",
       y = "Region") +
  theme_classic()
```

Oh wait, we haven't accounted for how ripe the banana is.... 

## Session details

```{r}
sessionInfo() |> 
  report::report()
```
